---
title: "eds221_day6_interactive"
author: "Charles Hendrickson"
date: "8/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
```

## Warm up

```{r}
rent_income <- us_rent_income %>% 
  select(-moe) %>% 
  pivot_wider(names_from = variable, values_from = estimate)

```

## Practice some tidyverse functions for data wrangling 

- `filter()`
- `select()`
- `mutate()`
- `case_when()`
- `group_by() %>%  summarize()`
- piping things together in sequence.


1. Exact match (single condition)

Only keep observations where the island is Biscoe
```{r}
ex_1 <- filter(.data = penguins, island == "Biscoe")
unique(ex_1$island)

```

# 2. "and" statement within filter

only keep observations if the penguin species is Chinstrap AND the year is 2008
```{r}
# This first option is the most common.
ex_2 <- penguins %>% 
  filter(species == "Chinstrap", year == 2008)

# Same thing as:

ex_2 <- penguins %>% 
  filter(species == "Chinstrap" & year == 2008)

# Also same thing as:

ex_2 <- penguins %>% 
  filter(species == "Chinstrap") %>%
  filter(year == 2008)

```

3. Exclude Gentoos

keep observations if the penguin species is NOT Gentoo
```{r}
ex_3 <- penguins %>% 
  filter(species != "Gentoo")
```

#Exercies 

#Create a subset that only contains Adelies and greater than 185 flipper length. 

```{r}

ex_3 <- penguins %>% 
  filter(species == "Adelie" & flipper_length_mm >= 185)

```

4. 
We want to keep observations where the island is Dream or torgersen.
```{r}
ex_4 <- penguins %>% 
  filter(island %in% c("Dream", "Torgersen"))



```

! = is not the 'island' value existing in the vector.
EX: 
```{r}
ex_4 <- penguins %>% 
  filter(!island %in% c("Dream", "Torgersen"))
```

### select() to select or exclude columns 

5. Keep a single column by name 

Only want to keep the species column.

```{r}
ex_5 <- penguins %>% select(species)

# rewrite this withou the pipe operator 

ex_5 <- select(.data = penguins, species)

```

6. Keep multiple columns by name 

create a sunset that contains species bill length_mm through flipper_length_mm, and year 
```{r}
ex_6 <- penguins %>% 
  select(species, bill_length_mm:flipper_length_mm, year)
```
 
 
7. Choose or exclude columns based on position
Select everything except for column 4. 
```{r}
ex_7 <- penguins %>% 
  select(-4)
```
 
8. A piped sequence 

Write a piped sequence where you first only keep columns for species, body_mass_g, and island, and second: filter to only include observations if body_mass_g is greater than 3600

```{r}
ex_8 <- penguins %>% 
  select(species, body_mass_g, island) %>% 
  filter(body_mass_g > 3600)
```

# mutate() to add or update existing columns, while keeping the ones already there. 

9. Add a column that has body mass converted to kg.


```{r}
ex_9 <- penguins %>% 
  mutate(body_mass_kg = body_mass_g / 1000) %>% 
  mutate(flipper_length_m = flipper_length_mm / 1000)

```

### `==` means "does this and this match"?



10. Mutate to update an existing column

```{r}
ex_10 <- penguins %>% 
  mutate(island_hooray = str_replace(island, pattern = "sen", 
                              replacement = "HOORAY"))
```

11. update the class of the bill depth column to character 
```{r}
ex_11 <- penguins %>% 
  mutate(bill_depth_mm = as.numeric(bill_depth_mm))

class(ex_11$bill_depth_mm)
```


### case_when(): your favorite if-else 

12. Create a new column that contains "small" if penguin flipper length is less than 185 mm, "medium" otherwise 

```{r}
ex_12 <- penguins %>% 
  mutate(binned_size = case_when(
  flipper_length_mm < 185 ~ "small",
  flipper_length_mm >= 185 ~ "medium"
  ))
```


```{r}
ex_13 <- penguins %>% 
  mutate(bird_bill_bin = case_when(
    bill_length_mm < 38 ~ "short bill",
    bill_length_mm >= 38 & bill_length_mm < 43 ~ "medium bill",
    bill_length_mm >= 43 ~ "medium bill", TRUE ~ "long bill"
  ))
```

### group_by() %>% summarize()

```{r}
# Mean of all bill_depth_mm values
#mean(penguins$bill_depth_mm, na.rm = TRUE)

# SO...

ex_14 <- penguins %>% 
  group_by(species, island, year, sex) %>% 
  summarize(mean_flipper_length = mean(flipper_length_mm, na.rm = TRUE),
            sd_flipper_length = sd(flipper_length_mm, na.rm = TRUE))



# Very different outcome: 
penguins %>% 
  group_by(species) %>% 
  mutate(mean_flipper_length = mean(flipper_length_mm, na.rm = TRUE)

```

# Pipe sequence practice 
```{r}
ex_15 <- penguins %>% 
  filter(body_mass_g > 3500) %>%
  select(species, flipper_length_mm, sex) %>% 
#group_by(sex, species)
  summarize(mean_flipper_length = mean(flipper_length_mm, na.rm = TRUE),
            sd_flipper_length = sd(flipper_length_mm, na.rm = TRUE),
            min_flipper_length = min(flipper_length_mm, na.rm = TRUE))





```

















